### 2020-10-15 paper/talk/talk.tex
Add beamer talk (no videos...)

\documentclass[slideopt,A4,showboxes,svgnames]{beamer}

%% list of packages here
\usepackage[absolute,showboxes,overlay]{textpos}
\usepackage{booktabs}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{pifont}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{array, makecell}
\usepackage{blkarray}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage[backend=biber, style=authoryear]{biblatex}
\renewcommand*{\nameyeardelim}{\addcomma\addspace}
\addbibresource{../references.bib}

\usepackage{theme/beamerthemeinria}
\input{mathdef}
\input{theme/style}

\title[Robust-Adaptive Control]{Robust-Adaptive Control\\of Linear Systems:\\beyond Quadratic Costs}
\subtitle{Sous-titre}
\date[date]{date}
\author[Edouard Leurent]{\textbf{Edouard Leurent}\inst{1,2},\\
	Odalric-Ambrym Maillard\inst{1},\\
	Denis Efimov\inst{1}}
\institute{
	\inst{1} Univ. Lille, Inria, CNRS, \\ ~Centrale Lille, UMR 9189 – CRIStAL,\\
	\inst{2} Renault Group}

\begin{document}

\begin{frame}
    \titlepage
\end{frame}

\begin{frame}{Motivation}
Most RL algorithms rely on trial and {\red error}
\begin{itemize}
	\item {\blue random} exploration
	\item {\green optimism} in the face of uncertainty
\end{itemize}
\leavevmode\newline
Our work: ensure safety at all time
\begin{itemize}
	\item {\red pessimism} in the face of uncertainty
	\item {\green learn} from observations
\end{itemize}
\end{frame}

\begin{frame}{Setting}
\begin{alertblock}{Dynamical priors with structured uncertainty}
	State $x$, control $u$, bounded disturbance $\omega$
	\begin{align*}
	\dot{x}(t) &= {\orange A(\theta)}x(t) + B u(t) + \omega(t),\\
	{\orange A(\theta)} &= A + \sum_{i=1}^d {\orange \theta_i}\phi_i,
	\end{align*}
	where $A,B,\phi_1,\dots,\phi_d$ are \alert{known}.
\end{alertblock}

\begin{alertblock}{Performance measure}
	\begin{itemize}
		\item Discounted return $\sum_n\gamma^n R(x(t_n))$.
		\item Reward $R$ only assumed to be \alert{bounded}.
	\end{itemize}
\end{alertblock}
\end{frame}

\begin{frame}{Robust control framework}
Given $N$ observations,
\begin{enumerate}
	\item Build a {\red confidence region} for the dynamics
	$$\probability{{\orange \theta}\in {\red \cC_{N,\delta}}} \geq 1-\delta,$$
	\item Maximize the {\red worst-case} performance ${\red V^r}$
	$${\green\sup_{\bu\in{(\Real^q)}^\Natural}} \underbrace{{\red\inf_{\substack{\theta \in \cC_{N,\delta} \\ \omega\in[\underline{\omega},\overline{\omega}]^\Real}}} \left[\sum_{n=N+1}^\infty \gamma^n R(x(t_n))\middle| {\green \bu},{\red \theta},{\red\omega}\right]}_{{\red V^r(\bu)}},$$
\end{enumerate}

Related work
\begin{itemize}
	\item Robust Dynamic Programming (finite states only)
	\item Linear-Quadratic setting (quadratic costs only)
\end{itemize}
\end{frame}

\begin{frame}{Algorithm 1/3}
\begin{block}{Model Estimation}
	Confidence ellipsoid from non-asymptotic linear regression.
	\[\| \theta_{N,\lambda}  - \theta\|_{G_{N,\lambda}} \leq \beta_N(\delta) \]
\end{block}

\begin{center}
	\includegraphics[width=0.4\linewidth]{../img/ellipsoid}
\end{center}

\end{frame}

\begin{frame}{Algorithm 2/3}
To optimise $V^r$ $\rightarrow$ approximate it by a tractable surrogate $\hat{V}^r$
\begin{block}{Interval Prediction} Propagate uncertainty ${\red \cC_{N,\delta}}$ through time and bound the state
	$${\blue \underline{x}(t)}\leq x(t)\leq{\blue \overline{x}(t)},\quad\forall t\geq t_N$$
\end{block}

\begin{center}
%	\includegraphics[width=0.6\linewidth]{../img/obstacle_small}
	\includegraphics[width=0.4\linewidth]{./img/interval-hull}
\end{center}

\end{frame}

\begin{frame}{Algorithm 3/3}

\begin{block}{Pessimistic Planning}
	Surrogate pessimistic return, maximized by tree based-planning
	\[{\orange \hat{V}^r(\bu)} \eqdef \sum_{n=N+1}^\infty \gamma^n \min_{x\in{\blue [\underline{x}(t_n), \overline{x}(t_n)]}}  R(x).\]
\end{block}
\begin{center}
	\includegraphics[trim={0 1.2cm 0 0}, clip, width=0.75\linewidth]{./img/planning}
\end{center}
\end{frame}

\begin{frame}{Results}
\begin{theorem}[Lower bound]
	\begin{equation*}
	{\orange \hat{V}^r(\bu)} \leq {\red V^r(\bu)} \leq V(u)
	\end{equation*}
\end{theorem}

\begin{theorem}[Suboptimality bound]
	Under two conditions:
	\begin{enumerate}
		\item Lipschitz rewards $R$:
		\item a stability condition over ${\red A(\theta_N)}$, for $N>N_0$
	\end{enumerate}
	we can bound the suboptimality with probability $1-\delta$ as:
	\begin{equation*}
	\underbrace{V(u_\star) - V(u_K)}_{\text{suboptimality}} \leq  {\red \underbrace{\Delta_\omega}_{\substack{\text{robustness to}\\ \text{disturbances}}}} + {\blue \underbrace{\cO\left(\frac{\beta_N(\delta)^2}{\lambda_{\min}(G_{N,\lambda})}\right)}_{\text{estimation error}}} + {\green \underbrace{\cO\left(K^{-\frac{\log 1/\gamma}{\log \kappa}}\right)}_{\text{planning error}}} 
	\end{equation*}
\end{theorem}
\end{frame}

\begin{frame}
\begin{corollary}[Asymptotic Near-Optimality]
	If the features $\phi_i x_n$ are \alert{persistently excited}, 
	\begin{itemize}
		\item the stability condition can be \alert{relaxed} to ${\orange A(\theta)}$ only
		\item the suboptimality bound takes the more explicit form
	\end{itemize}
	\begin{equation*}
	{V(u_\star) - V(u_K)}\leq  {\red \Delta_\omega} + {\blue {\cO\left(\frac{\log(N^{d/2}/\delta)}{N}\right)}} + {\green {\cO\left(K^{-\frac{\log 1/\gamma}{\log \kappa}}\right)}}
	\end{equation*}
	which ensures near-optimality as ${\blue N\to\infty}$ and ${\green K\to\infty}$.
\end{corollary}
\end{frame}

\begin{frame}{Experiments}

\end{frame}

\begin{frame}
\centering \LARGE Thank You!\\[1cm]
\large \emph{I am looking for a postdoctoral position.}
\end{frame}

\end{document}